---
title: "README.tmp000"
output: html_document
---

## Running the project

The simplest (and recommended) way to run the project for the very first time is to run the [__run-all.R](r/_startup/__run-all.R) script located in the [_startup](r/_startup) folder.

This script is a runner for the [_starter.R](r/_startup/_starter.R) script, which in turn is a launcher for the following four scripts:

1. [0.setup.R](r/_startup/0.setup.R): The script is a launcher for the [setup.R](r/src/support-scripts/setup.R) script mentioned above in the [Essential Directories and Files] section.
2. [1.init.R](r/_startup/1.init.R): Initializes all the required datasets and loads the general-purpose and data helper functions from the following scripts:
  - [logging-functions.R](r/src/support-functions/logging-functions.R): User-defined *logging functions*;
  - [common-helper.functions.R](r/src/support-functions/common-helper.functions.R): User-defined *common helper functions*;
  - [data.helper.functions.R](r/src/support-functions/data.helper.functions.R): User-defined *data helper functions*.
3.[2.run-main-script.R](r/_startup/2.run-main-script.R): The script is a launcher for the [capstone-movielens.main.R](r/src/capstone-movielens.main.R) script mentioned above in the [Essential Directories and Files] section.
4.[3.rmd.render.R](r/_startup/3.rmd.render.R): Renders the *Capstone PDF Report*.
  
> [!NOTE]
> The very first execution of the script [__run-all.R](r/_startup/__run-all.R) may take several hours.

Subsequent executions of the project scripts will be significantly faster, as the data stored in the cache during the initial run will not be recalculated.

To force recalculation of the particular data, the corresponding cache file must be manually deleted. 

Additionally, project scripts can subsequently be run individually to perform a specific task.
These tasks are described in more detail below.

Before running the scripts below, make sure that the following variables, containing the *File System paths*, are initialized:

- `startup.path`;
- `setup.script_path`;
- `init.script_path`;
- `run_main.script_path`;
- `rmd_render.script_path`;

These variables can be initialized by the following [code snippet](r/_startup/_starter.R#L2) provided in the script [_starter.R](r/_startup/_starter.R#L2):
```{r starter-script@l2, eval=FALSE}
## Source File Paths -----------------------------------------------------------
startup.path <- "r/_startup"
setup.script_path <- file.path(startup.path, "0.setup.R")
init.script_path <- file.path(startup.path, "1.init.R")
run_main.script_path <- file.path(startup.path, "2.run-main-script.R")
rmd_render.script_path <- file.path(startup.path, "3.rmd.render.R")
```

### Installing Packages and/or Loading External Libraries

As mentioned above in section [Essential Directories and Files], the script [setup.R](r/src/support-scripts/setup.R) installs all necessary packages, loads required libraries, and resolves conflicts.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L11) provided in the script [_starter.R](r/_startup/_starter.R#L11):
```{r starter-script@l11, eval=FALSE}
### Installing Packages and/or Loading External Libraries
source(setup.script_path, 
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Initializing Project Datasets and Loading Core Helper Functions

As mentioned above in *this section*, the script [1.init.R](r/_startup/1.init.R) initializes all the required datasets and loads the general-purpose and data helper functions.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L19) provided in the script [_starter.R](r/_startup/_starter.R#L19):
```{r starter-script@l19, eval=FALSE}
### Initializing Project Datasets and Loading Core Helper Functions
source(init.script_path, 
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Launching the *Main `R` Script*

As mentioned above in section [Essential Directories and Files], the script [capstone-movielens.main.R](r/src/capstone-movielens.main.R) is the *Project Main `R` Script*.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L27) provided in the script [_starter.R](r/_startup/_starter.R#L27):
```{r starter-script@l27, eval=FALSE}
### Launching the *Main `R` Script*
source(run_main.script_path,
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Rendering the Capstone PDF Report

As mentioned above in *this section*, the script [3.rmd.render.R](r/_startup/3.rmd.render.R) renders the *Capstone PDF Report*.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L35) provided in the script [_starter.R](r/_startup/_starter.R#L35):
```{r starter-script@l35, eval=FALSE}
### Rendering the Capstone PDF Report
source(rmd_render.script_path,
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```
