---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# edx.Capstone Choose Your Own Project

*A movie recommendation system using the MovieLens dataset*

This is the _**Capstone Choose Your Own**_ Project as part of the [edX HarvardX PH125.9x Data Science Capstone Course](https://pll.harvard.edu/course/data-science-capstone) to build a _Choose Your Own Project_ according to the course requirements.

> [!NOTE]
> This project is implemented as [RStudio Project](https://support.posit.co/hc/en-us/articles/200526207-Using-RStudio-Projects).

The goal of the project is to build a _Handwritten Character Recognition System_ using the [HandWritten_Character Dataset](https://www.kaggle.com/datasets/vaibhao/handwritten-characters)  downloaded from the  [Kaggle](https://www.kaggle.com) platform.

As described on the [landing web-page], the *context* of the dataset is as follows:

1. Used *EMNIST* data for Alphabets and Digits.
2. Transformed the Data using some image processing techniques and convert it to 32,32 pixel black and white images.
3. Created the Data set for special character ( @, #, $, & )
3. Merged the Categories to avoid the misclassification
4. Total 39 Categories in Train and Validation set

## Essential Directories and Files

The following are essential directories and files of the project:

- [edx.capstone.choose-your-own.Rproj](edx.capstone.choose-your-own.Rproj): The [RStudio Project](https://support.posit.co/hc/en-us/articles/200526207-Using-RStudio-Projects) file;

- [index.R](r/index.R): The *Index (main) `R` Script*;

- [support-functions](r/support-functions): Folder for the *User-defined Functions* used throughout the project;

- [support-scripts](r/support-scripts): Folder for the *Auxiliary (Support) Scripts* used by the *Index `R` Script*;

- [setup.R](r/support-scripts/setup.R): Support script for installing all necessary packages, loading required libraries, and resolving conflicts;

- [prepare-input-data.R](r/support-scripts/.R): Support script

- [.R](r/support-scripts/.R): Support script

- [HW-Chars.Recognition.Site](reports/HW-Chars.Recognition.Site): Folder for the *RMD-Report* files;

- [capstone.choose-your-own.report.pdf](#): (Not implemented yet) The final *PDF report* - the ultimate product of the project;

- `data` folder: (automatically created if it does not exist, after the first project run) Directory for all data files used in the project;  
  
## Working Environment

Since this project is an [RStudio Project](https://support.posit.co/hc/en-us/articles/200526207-Using-RStudio-Projects), it is recommended to work with the project in *RStudio IDE*.
As mentioned above in section [Essential Directories and Files], the *RStudio Project* file is [Capstone-MovieLens.Rproj](Capstone-MovieLens.Rproj).

## Running the project

The simplest (and recommended) way to run the project for the very first time is to run the [__run-all.R](r/_startup/__run-all.R) script located in the [_startup](r/_startup) folder.

This script is a runner for the [_starter.R](r/_startup/_starter.R) script, which in turn is a launcher for the following four scripts:

1. [0.setup.R](r/_startup/0.setup.R): The script is a launcher for the [setup.R](r/src/support-scripts/setup.R) script mentioned above in the [Essential Directories and Files] section.
2. [1.init.R](r/_startup/1.init.R): Initializes all the required datasets and loads the general-purpose and data helper functions from the following scripts:
  - [logging-functions.R](r/src/support-functions/logging-functions.R): User-defined *logging functions*;
  - [common-helper.functions.R](r/src/support-functions/common-helper.functions.R): User-defined *common helper functions*;
  - [data.helper.functions.R](r/src/support-functions/data.helper.functions.R): User-defined *data helper functions*.
3.[2.run-main-script.R](r/_startup/2.run-main-script.R): The script is a launcher for the [capstone-movielens.main.R](r/src/capstone-movielens.main.R) script mentioned above in the [Essential Directories and Files] section.
4.[3.rmd.render.R](r/_startup/3.rmd.render.R): Renders the *Capstone PDF Report*.
  
> [!NOTE]
> The very first execution of the script [__run-all.R](r/_startup/__run-all.R) may take several hours.

Subsequent executions of the project scripts will be significantly faster, as the data stored in the cache during the initial run will not be recalculated.

To force recalculation of the particular data, the corresponding cache file must be manually deleted. 

Additionally, project scripts can subsequently be run individually to perform a specific task.
These tasks are described in more detail below.

Before running the scripts below, make sure that the following variables, containing the *File System paths*, are initialized:

- `startup.path`;
- `setup.script_path`;
- `init.script_path`;
- `run_main.script_path`;
- `rmd_render.script_path`;

These variables can be initialized by the following [code snippet](r/_startup/_starter.R#L2) provided in the script [_starter.R](r/_startup/_starter.R#L2):
```{r starter-script@l2, eval=FALSE}
## Source File Paths -----------------------------------------------------------
startup.path <- "r/_startup"
setup.script_path <- file.path(startup.path, "0.setup.R")
init.script_path <- file.path(startup.path, "1.init.R")
run_main.script_path <- file.path(startup.path, "2.run-main-script.R")
rmd_render.script_path <- file.path(startup.path, "3.rmd.render.R")
```

### Installing Packages and/or Loading External Libraries

As mentioned above in section [Essential Directories and Files], the script [setup.R](r/src/support-scripts/setup.R) installs all necessary packages, loads required libraries, and resolves conflicts.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L11) provided in the script [_starter.R](r/_startup/_starter.R#L11):
```{r starter-script@l11, eval=FALSE}
### Installing Packages and/or Loading External Libraries
source(setup.script_path, 
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Initializing Project Datasets and Loading Core Helper Functions

As mentioned above in *this section*, the script [1.init.R](r/_startup/1.init.R) initializes all the required datasets and loads the general-purpose and data helper functions.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L19) provided in the script [_starter.R](r/_startup/_starter.R#L19):
```{r starter-script@l19, eval=FALSE}
### Initializing Project Datasets and Loading Core Helper Functions
source(init.script_path, 
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Launching the *Main `R` Script*

As mentioned above in section [Essential Directories and Files], the script [capstone-movielens.main.R](r/src/capstone-movielens.main.R) is the *Project Main `R` Script*.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L27) provided in the script [_starter.R](r/_startup/_starter.R#L27):
```{r starter-script@l27, eval=FALSE}
### Launching the *Main `R` Script*
source(run_main.script_path,
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```

### Rendering the Capstone PDF Report

As mentioned above in *this section*, the script [3.rmd.render.R](r/_startup/3.rmd.render.R) renders the *Capstone PDF Report*.
It can also be launched by the following [code snippet](r/_startup/_starter.R#L35) provided in the script [_starter.R](r/_startup/_starter.R#L35):
```{r starter-script@l35, eval=FALSE}
### Rendering the Capstone PDF Report
source(rmd_render.script_path,
       catch.aborts = TRUE,
       echo = TRUE,
       spaced = TRUE,
       verbose = TRUE,
       keep.source = TRUE)
```
